---
title: "TS model diagnostics"
output: github_document
---

Summary report of various diagnostics for LDATS time series models run with model configurations.

The current diagnostics are:

* Runtime (in seconds)
* AIC as reported in the `TS_on_LDA` object from LDATS; `ts_model$AIC`
* AICc as calculated by `LDATS::AICc`; `LDATS::AICc(ts_model)`
* Trace plots of parameter estimates over time. These parameters are:
  * Etas: Estimates of the intercept & covariate coefficients (if applicable) for the TS fit. For a model with `n` changepoints and `k` topics, there will be `(n + 1) * (k - 1)` etas. 
  * Rhos: Estimates of the changepoint locations (if applicable). There is one rho for every changepoint in a model. 

The current model configurations are all combinations of:
* LDA seed = 1977
* 2, 3, 6, or 12 topics
* 0, 1, 2, 3, or 4 changepoints
* Formulas `~ 1` or `~ time`

I have not plotted all the etas because there are a lot of them for the more complex models. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(drake)
library(dplyr)
library(ggplot2)

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake", "drake-cache-manyoptions.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)

```

```{r load stuff}
loadd(all_models, cache = cache)

model_names <- lapply(as.list(names(all_models)), 
                      FUN = function(model_result_name)
                        return(strsplit(model_result_name, "_")[[1]]))

names(model_names) = 1:length(model_names)
model_names <- bind_rows(model_names) %>%
  t()

model_names <- as.data.frame(model_names) %>%
  select(-V1) %>%
  rename(seed = V2,
         k = V3,
         ncpts = V4,
         cov = V5,
         nit = V6) %>%
  mutate(k = factor(k),
         ncpts = as.factor(ncpts),
         obj_index = row_number(),
         full_name = names(all_models)) %>%
  mutate(k = factor(k, levels = as.character(sort(as.numeric(levels(k))))))

```


### Runtime of TS models (in seconds)

The y-axis is runtime; colors are the number of topics; the x axis is the number of iterations. The facets are the number of changepoints (columns) and the formula (rows).

More changepoints takes longer, and of course more iterations takes longer.

```{r runtime}
runtime <- data.frame(
  full_name = names(all_models),
  runtime = vapply(all_models,
                   FUN = function(model_result)
                     return(model_result$timing$toc["elapsed"] -
                              model_result$timing$tic["elapsed"]),
                   FUN.VALUE = 8),
  row.names = NULL, stringsAsFactors = F)

model_info <- left_join(model_names, runtime, by = "full_name")


runtime_plot <-ggplot(data = model_info, aes(x = nit, y = runtime, color = k)) +
geom_boxplot() +
  facet_grid(rows = vars(cov), cols = vars(ncpts), switch = "y") +
  theme_bw() + 
  ggtitle("Runtime") +
    scale_color_viridis_d(end = .8)

runtime_plot

```

## AIC and AICc of TS models

```{r aicc}
aiccs <- data.frame(
  aicc = vapply(all_models,
                FUN = function(ts_result)
                  return(LDATS::AICc(ts_result$ts[[1]])),
                FUN.VALUE = 30),
  aic = vapply(all_models,
               FUN = function(ts_result)
                 return(ts_result$ts[[1]]$AIC),
               FUN.VALUE = 30),
  full_name = names(all_models),
  row.names = NULL, stringsAsFactors = F)

model_info <- left_join(model_info, aiccs, by = "full_name")

aic_plot <- ggplot(data = model_info, aes(x = nit, y = aic, color = k)) +
  geom_boxplot() +
  facet_grid(rows = vars(cov), cols = vars(ncpts), switch = "y") +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  ggtitle("AIC")

aic_plot

aic_match <- model_info %>%
  group_by(ncpts, k, cov) %>%
  arrange(.by_group = T) %>%
  ungroup() %>%
  select(-seed, -obj_index, -full_name, -runtime)


```


```{r aiccs}


aicc_plot <- ggplot(data = model_info, aes(x = nit, y = aicc, color = k)) +
  geom_boxplot() +
  facet_grid(rows = vars(cov), cols = vars(ncpts), switch = "y") +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  ggtitle("AICc")

aicc_plot

```

AIC appears not to change across model configurations (i.e. between `~intercept` and `~time`), but the values are in fact different. The values *do not* change across numbers of iterations, however.

```{r print aic match}
print(aic_match[1:30, ])
```


## Parameter estimates over iterations

### Etas (coefficients within segments)

I have plotted etas for models with 2 topics, 0 or 1 changepoint, and `~1` or `~time`. Other configurations are possible but can be very large, because there are new etas for every additional topic-segment combination. I'm hoping the 10000 iteration plots will be visible at high resolution.

The facet strips are: number of changepoints; formula; parameter being estimated.

```{r etas}

get_etas <- function(ts_result) {
  etas_df <- ts_result$ts[[1]]$etas %>%
    as.data.frame() %>%
    mutate(draw = row_number())
  return(etas_df)
}

etas <- lapply(all_models, FUN = get_etas)

etas <- bind_rows(etas, .id = "full_name") %>%
  tidyr::gather(-draw, -full_name, key = "parameter", value = "estimate") %>%
  filter(!is.na(estimate))

etas_info <- left_join(etas, model_info, by = "full_name") %>%
  filter(as.character(k) %in% c("2"),
         as.character(ncpts) %in% c("0L", "1L")) 


etas_plots <- list() 

for(i in 1:length(levels(etas_info$nit))) {
  these_etas <- filter(etas_info, nit == levels(nit)[i])
  
  lsize <- max(.05, 100 / as.numeric(as.character(these_etas$nit[1])))
  
  etas_plots[[i]] <- ggplot(data = these_etas, aes(x = draw, y = estimate)) +
  geom_line(size = lsize) +
  theme_bw() +
  facet_wrap(facets = c("ncpts", "cov", "parameter"), scales = "free", strip.position = "top", ncol = 1, drop = TRUE)  +
  scale_color_viridis_d(end = .8) +
    ggtitle(paste0("Etas: ", these_etas$nit[1], " iterations"))
  
}

```

#### Etas 100 iterations

```{r etas 100, fig.width = 10, fig.height = 20}

etas_plots[[1]]

```

#### Etas 1000 iterations

```{r etas 1000, fig.width = 10, fig.height = 20}

etas_plots[[2]]

```

#### Etas 10000 iterations

```{r etas 10000, fig.width = 25, fig.height = 20}

etas_plots[[3]]

```

### Rhos (changepoint locations)

I have plotted rhos for models with 2 topics, 1 or 2 changepoints, and `~1` or `~time`. The facet strips are number of changepoints; formula. The colors are the different changepoints being estimated.

```{r rhos}

get_rhos <- function(ts_result) {
  rhos_df <- ts_result$ts[[1]]$rhos %>%
    as.data.frame() %>%
    mutate(draw = row_number())
  return(rhos_df)
}

rhos <- lapply(all_models, FUN = get_rhos)

rhos <- bind_rows(rhos, .id = "full_name") %>%
  tidyr::gather(-draw, -full_name, key = "changepoint", value = "estimate") %>%
  filter(!is.na(estimate)) %>%
  mutate(changepoint = substr(changepoint, 2, nchar(changepoint)))

rhos_info <- left_join(rhos, model_info, by = "full_name") %>%
  filter(as.character(k) %in% c("2"),
         as.character(ncpts) %in% c("1L", "2L"))


rhos_plots <- list() 

for(i in 1:length(levels(rhos_info$nit))) {
  these_rhos <- filter(rhos_info, nit == levels(nit)[i])
  
  lsize <- max(.05, 100 / as.numeric(as.character(these_rhos$nit[1])))
  
  rhos_plots[[i]] <- ggplot(data = these_rhos, aes(x = draw, y = estimate, color = changepoint)) +
  geom_line(size = lsize) +
  theme_bw() +
  facet_wrap(facets = c("ncpts", "cov"), scales = "free", strip.position = "top", ncol = 1, drop = TRUE)  +
  scale_color_viridis_d(end = .8) +
    ggtitle(paste0("Rhos: ", these_rhos$nit[1], " iterations"))
  
}

```

#### Rhos 100 iterations

```{r rhos 100, fig.width = 10, fig.height = 20}

rhos_plots[[1]]

```

#### Rhos 1000 iterations

```{r rhos 1000, fig.width = 10, fig.height = 10}

rhos_plots[[2]]

```

#### Rhos 10000 iterations

```{r rhos 10000, fig.width = 25, fig.height = 10}

rhos_plots[[3]]

```