---
title: "TS model diagnostics"
output: github_document
---

Summary report of various diagnostics for LDATS time series models run with model configurations.

The current diagnostics are:

* Runtime (in seconds)
* AIC as reported in the `TS_on_LDA` object from LDATS; `ts_model$AIC`
* AICc as calculated by `LDATS::AICc`; `LDATS::AICc(ts_model)`
* Trace plots of parameter estimates over time. These parameters are:
  * Etas: Estimates of the intercept & covariate coefficients (if applicable) for the TS fit. For a model with `n` changepoints and `k` topics, there will be `(n + 1) * (k - 1)` etas. 
  * Rhos: Estimates of the changepoint locations (if applicable). There is one rho for every changepoint in a model. 

The current model configurations are all combinations of:
* LDA seed = 1977
* 2 or 5 topics
* 0, 1, or 4 changepoints
* Formulas `~ 1` or `~ time`
* 100, 1000, or 10000 iterations
* `penultimate_temp` = 2^4, 2^6, 2^8 (`LDATS` default is 2^6)

I have not plotted all the etas because there are a lot of them for the more complex models. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(drake)
library(dplyr)
library(ggplot2)

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake", "drake-cache.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)

```

```{r load stuff}
loadd(all_models, cache = cache)

model_names <- lapply(as.list(names(all_models)), 
                      FUN = function(model_result_name)
                        return(strsplit(model_result_name, "_")[[1]]))

names(model_names) = 1:length(model_names)
model_names <- bind_rows(model_names) %>%
  t()

model_names <- as.data.frame(model_names) %>%
  select(-V1) %>%
  rename(seed = V2,
         k = V3,
         ncpts = V4,
         cov = V5,
         nit = V6,
         penult_temp = V7) %>%
  mutate(k = factor(k),
         ncpts = as.factor(ncpts),
         penult_temp = as.factor(penult_temp),
         obj_index = row_number(),
         full_name = names(all_models)) %>%
  mutate(k = factor(k, levels = as.character(sort(as.numeric(levels(k))))))

```


### Runtime of TS models (in seconds)

The y-axis is runtime. The x axis is number of changepoints. The facet columns are the penultimate temperature, and the facet rows are covariate nested within number of topics.

More changepoints takes longer, and of course more iterations takes longer. The models speed up as the number of iterations increases (that is, 100000 iterations does not take 10x as long as 10000 iterations). The number of topics, penultimate temperature, and covariate don't impact runtime nearly as much as nit and ncpts. 

```{r runtime}
runtime <- data.frame(
  full_name = names(all_models),
  runtime = vapply(all_models,
                   FUN = function(model_result)
                     return(model_result$timing$toc["elapsed"] -
                              model_result$timing$tic["elapsed"]),
                   FUN.VALUE = 8),
  row.names = NULL, stringsAsFactors = F)

model_info <- left_join(model_names, runtime, by = "full_name")


runtime_plot <-ggplot(data = model_info, aes(x = ncpts, y = runtime, color = penult_temp)) +
geom_boxplot() +
  facet_grid(rows = vars(k, cov), cols = vars(nit), switch = "y") +
  theme_bw() + 
  ggtitle("Runtime") +
    scale_color_viridis_d(end = .8)

runtime_plot

```

## AIC and AICc of TS models

```{r aicc}
aiccs <- data.frame(
  aicc = vapply(all_models,
                FUN = function(ts_result)
                  return(LDATS::AICc(ts_result$ts[[1]])),
                FUN.VALUE = 30),
  aic = vapply(all_models,
               FUN = function(ts_result)
                 return(ts_result$ts[[1]]$AIC),
               FUN.VALUE = 30),
  full_name = names(all_models),
  row.names = NULL, stringsAsFactors = F)

model_info <- left_join(model_info, aiccs, by = "full_name")

aic_plot <- ggplot(data = model_info, aes(x = nit, y = aic, color = k)) +
  geom_boxplot() +
  facet_grid(rows = vars(cov), cols = vars(ncpts, penult_temp), switch = "y") +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  ggtitle("AIC")

aic_plot

```

The AICcs are suspect right now.


```{r aiccs, eval = F}

aicc_plot <- ggplot(data = model_info, aes(x = nit, y = aicc, color = k)) +
  geom_boxplot() +
  facet_grid(rows = vars(cov), cols = vars(ncpts, penult_temp), switch = "y") +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  ggtitle("AICc")

aicc_plot

```

```{r aic matched}

aic_match <- model_info %>%
  group_by(ncpts, k, cov, penult_temp) %>%
  arrange(.by_group = T) %>%
  ungroup() %>%
  select(-seed, -obj_index, -full_name, -runtime)

print(aic_match[1:10, ])

print(aic_match[37:47, ])

```

AIC and AICc do not change, out to 7 figures, over iterations or temperature for 0 changepoint models. For models with changepoints, AIC and AICc change numerically but generally negligably (<2 units). 

## Parameter estimates over iterations

### Etas (coefficients within segments)


~I have plotted etas for models with 2 topics, 0 or 1 changepoint, and `~1` or `~time`. Other configurations are possible but can be very large, because there are new etas for every additional topic-segment combination. I'm hoping the 10000 iteration plots will be visible at high resolution.~

~The facet strips are: number of changepoints; formula; temperature; parameter being estimated.~

Now I have plotted the mean and SD of etas estimates, split into and calculated within segments of 100 draws. The y axis is the mean or sd estimate, and the x axis is the draw (/100). Color is the nb of iterations, and the facet strips are ncpts, cov, temp, and parameter. 


```{r etas, eval = T}

get_etas <- function(ts_result) {
  etas_df <- ts_result$ts[[1]]$etas %>%
    as.data.frame() %>%
    mutate(draw = row_number()) %>%
    mutate(draw_100th = ceiling(draw / 100))
  return(etas_df)
}

etas <- lapply(all_models, FUN = get_etas)

etas <- bind_rows(etas, .id = "full_name") %>%
  tidyr::gather(-draw, -draw_100th, -full_name, key = "parameter", value = "estimate") %>%
  filter(!is.na(estimate)) %>%
  group_by(full_name, draw_100th, parameter) %>%
  mutate(mean_100th = mean(estimate),
         sd_100th = sd(estimate)) %>%
  ungroup()

etas_info <- left_join(etas, model_info, by = "full_name") %>%
  filter(as.character(k) %in% c("2"),
         as.character(ncpts) %in% c("0", "1")) 


etas_plots <- list() 

for(i in 1:length(levels(etas_info$nit))) {
  these_etas <- filter(etas_info, nit == levels(nit)[i])
  

  if(i < 3) {
  lsize <- max(.05, 100 / as.numeric(as.character(these_etas$nit[1])))
  } else {
    lsize <- .05
  }
    
  etas_plots[[i]] <- ggplot(data = these_etas, aes(x = draw, y = estimate)) +
  geom_line(size = lsize) +
  theme_bw() +
  facet_wrap(facets = c("ncpts", "cov", "penult_temp", "parameter"), scales = "free", strip.position = "top", ncol = 1, drop = TRUE)  +
  scale_color_viridis_d(end = .8) +
    ggtitle(paste0("Etas: ", these_etas$nit[1], " iterations"))
  
}


etas_100ths <-  unique(select(etas_info, ncpts, cov, penult_temp, parameter, nit, full_name, draw_100th, mean_100th, sd_100th))

etas_means_100ths <- ggplot(data =etas_100ths, aes(x = draw_100th, y = mean_100th, color = nit)) +
  facet_wrap(facets = c("ncpts", "cov", "penult_temp", "parameter"), ncol = 3, scales = "free") +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  ggtitle("Mean of parameter estimates")
```

#### Mean plots
```{r eta means plots, fig.width = 24, fig.height = 48}
etas_means_100ths

```

#### SD plots

```{r eta sd plots, fig.width = 24, fig.height = 48}
etas_sd_100ths <- ggplot(data =etas_100ths, aes(x = draw_100th, y = sd_100th, color = nit)) +
  facet_wrap(facets = c("ncpts", "cov", "penult_temp", "parameter"), ncol = 3, scales = "free") +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  ggtitle("SD of parameter estimates")

etas_sd_100ths

```


#### Etas 100 iterations

```{r etas 100, fig.width = 10, fig.height = 60, eval = F}

etas_plots[[1]]

```


#### Etas 1000 iterations

```{r etas 1k, fig.width = 20, fig.height = 60, eval = F}

etas_plots[[2]]

```


#### Etas 100,000 iterations

```{r etas 100k, fig.width = 40, fig.height = 60, eval= F}

etas_plots[[3]]

```

### Rhos (changepoint locations)

I have plotted rhos for models with 2 topics, 1 or 4 changepoints, and `~1` or `~time`. The facet strips are number of changepoints; temperature; formula. The colors are the different changepoints being estimated.

```{r rhos, eval = F}

get_rhos <- function(ts_result) {
  rhos_df <- ts_result$ts[[1]]$rhos %>%
    as.data.frame() %>%
    mutate(draw = row_number())
  return(rhos_df)
}

rhos <- lapply(all_models, FUN = get_rhos)

rhos <- bind_rows(rhos, .id = "full_name") %>%
  tidyr::gather(-draw, -full_name, key = "changepoint", value = "estimate") %>%
  filter(!is.na(estimate)) %>%
  mutate(changepoint = substr(changepoint, 2, nchar(changepoint)))

rhos_info <- left_join(rhos, model_info, by = "full_name") %>%
  filter(as.character(k) %in% c("2"),
         as.character(ncpts) %in% c("1", "4"))


rhos_plots <- list() 

for(i in 1:length(levels(rhos_info$nit))) {
  these_rhos <- filter(rhos_info, nit == levels(nit)[i])
  
  if(i < 3) {
  lsize <- max(.05, 100 / as.numeric(as.character(these_rhos$nit[1])))
  } else {
    lsize <- .05
  }
  
  
  rhos_plots[[i]] <- ggplot(data = these_rhos, aes(x = draw, y = estimate, color = changepoint)) +
  geom_line(size = lsize) +
  theme_bw() +
  facet_wrap(facets = c("ncpts", "cov", "penult_temp"), scales = "free", strip.position = "top", ncol = 1, drop = TRUE)  +
  scale_color_viridis_d(end = .8) +
    ggtitle(paste0("Rhos: ", these_rhos$nit[1], " iterations"))
  
}

```

#### Rhos 100 iterations

```{r rhos 100, fig.width = 10, fig.height = 60, eval = F}

rhos_plots[[1]]

```

#### Rhos 1000 iterations

```{r rhos 1000, fig.width = 20, fig.height = 60, eval = F}

rhos_plots[[2]]

```

#### Rhos 10,000 iterations

```{r rhos 10000, fig.width = 20, fig.height = 60, eval = F}

rhos_plots[[3]]

```